Delocalization of Eigenvectors in Random Matrices

### 2025-03-24 Draft.tex
deloc left

\begin{boxtheorem}{Theorem 1: (Stopping time).} $\mathbb{P}[\tau_{\text{stop}, i} \neq  1] \lesssim_D N^{-D}$ for any $i=1, 2$ and $D > 0$.
\end{boxtheorem}
\textit{Proof} - By Boole's inequality: $$\mathbb{P}\left[\tau_{\sto, 1}\neq 1\right]\leq \mathbb{P}[\left\{\tau_{\sto, 2}\neq 1\}\cap \{\tau_{\sto, 1=1}\}\right]+P[\tau_\sto=\tau_{\sto, 1}<1]$$
On the event of the left term, we have $|\E_{\tau_\sto}(z)_{ab}|\geq W^{\delta_\sto}W^{-3/4}\eta_t^{-1}W^{-1}\eta_t^{-1/2}$ by our ear

\subsection*{7.2 Theorem 2 - Quantum Diffusion}
\begin{boxtheorem}{Theorem 2: (Quantum Diffusion)}
For $|E|<2$ fixed, assume $\exists \nu>0$, s.t. $\eta \asymp W^2 N^{-2}$ and $W\geq W^{8/11+\nu}$. Then: $$\max_{x, y}|T_{xy}-\Theta_{xy}|\prec W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$$
\end{boxtheorem}
\textit{Proof} -

@@

\noindent Now, let $\epsilon, D>0$ be arbitrary. Given that we work in the context of the delocalization, we have $W\gg N^{\frac{1}{2}+\beta} = N^{\frac{1+2\beta}{2}}\Rightarrow W^{-1/2}\ll N^{-\frac{1+2\beta}{4}}$. Hence, let us choose $\delta = \min \left(1, \frac{2\beta+1}{4\epsilon}\right)$. This is to ensure that $\delta\epsilon<\frac{2\beta+1}{4}$. By Def.1 of stochastic domination, from $(8.7)$, we have that: $$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)>N^{\delta\epsilon}(1+W^{-1/2}\Xi_N(s))\right)<N^{-D}$$
$$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)>N^{\delta\epsilon}+N^{\delta\epsilon} W^{-1/2}\Xi_N(s)\right)<N^{-D}$$
$$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)(1-N^{\delta\epsilon} W^{-1/2})>N^{\delta\epsilon}\right)<N^{-D}$$
$$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)>\frac{N^{\delta\epsilon}}{(1-N^{\delta\epsilon} W^{-1/2})}\right)<N^{-D}$$
By the delocalization scale, we have that $$N^{\delta\epsilon}W^{-1/2} \ll N^{\delta\varepsilon} \cdot N^{-1/4-\beta/2} = N^{\delta\varepsilon-1/4-\beta/2}$$
$$1-N^{\delta\epsilon} W^{-1/2}\gg 1-N^{\delta\epsilon -\frac{1+2\beta}{4}}$$
$$\frac{N^{\delta\epsilon}}{1-N^{\delta\epsilon} W^{-1/2}}\ll \frac{N^{\delta\epsilon}}{1-N^{{\delta\epsilon}-\frac{1+2\beta}{4}}}$$

\noindent This means that $\exists N_0$, s.t. for all $N\geq N_0$, the latter event is a subset: $$ \left\{\Xi_N(s)>\frac{N^\epsilon}{1-N^{{\delta\epsilon}-\frac{1+2\beta}{4}}}\right\}\subset \left\{\Xi_N(s)>\frac{N^{\delta\epsilon}}{(1-N^{\delta\epsilon} W^{-1/2})}\right\}$$
$$\mathbb{P} \left\{\Xi_N(s)>\frac{N^{\delta\epsilon}}{1-N^{\delta\epsilon-\frac{1+2\beta}{4}}}\right\}\leq\mathbb{P} \left\{\Xi_N(s)>\frac{N^{\delta\epsilon}}{(1-N^{\delta\epsilon} W^{-1/2})}\right\}\leq N^{-D}$$
 \textcolor{red}{COMPLETE THE STOCHASTIC DOMINATION}. More specifically, find $N^\gamma$

>>>>

\begin{boxtheorem}{Theorem 1: (Stopping time).} $\mathbb{P}[\tau_{\text{stop}, i} \neq  1] \lesssim_D N^{-D}$ for any $i=1, 2$ and $D > 0$.
\end{boxtheorem}
\textit{Proof} - By Boole's inequality: $$\mathbb{P}\left[\tau_{\sto, 1}\neq 1\right]\leq \mathbb{P}[\left\{\tau_{\sto, 2}\neq 1\}\cap \{\tau_{\sto, 1=1}\}\right]+\mathbb{P}[\tau_\sto=\tau_{\sto, 1}<1]$$
For any $\mathcal{A}\in \{\tau_\sto = \tau_{\sto, 1}<1\}$, we have $|\E_{\tau_\sto}(z)_{ab}|\geq W^{\delta_\sto}W^{-3/4}\eta_t^{-1}W^{-1}\eta_t^{-1/2}$, since $\E_t(z)$ is almost surely continuous in $t$. By definition of the stopped SDE, we have $\E_{\tau_\sto}(z) = \E_{\tau_\sto}^{\sto}(z)$, i.e: $$\mathbb{P}\left[\tau_\sto = \tau_{\sto, 1}<1\right]\leq \mathbb{P}\left[\max_{t\in [0, 1]}\max_{a, b}\frac{|\E_t^{M, \sto}(z)_{ab}|}{W^{-\frac{3}{4}}\eta_t^{-\frac{3}{4}}W^{-1}\eta_t^{-\frac{1}{2}}}\geq W^{\delta_\sto}\right]$$
By our theorems $6.1$ through $6.3$ in the last section applied to the RHS, we know from the definition of stochastic domination that for any $D>0$: 
$$\mathbb{P}\left[ \tau_\sto = \tau_{\sto, 1}<1\right]\leq C_DN^{-D}$$
This means that we know need to establish an equivalent bound for the first term on the RHS. Assume an event $\mathcal{B}$ for which the following is true:
$$\max_{t\in [0, 1]}\max_{a, b}\frac{|\E_t(z)_{ab}|}{W^{\delta_\sto}W^{-\frac{3}{4}}\eta_t^{-\frac{3}{4}}W^{-1}\eta_t^{-\frac{1}{2}}}\lesssim 1$$
Since $T_t(z) = \Theta_t+\E_t(z)$ and $|(\Theta_t)_{ab}|\lesssim W^{-1}\eta_t^{-1/2}$ (Lemma A.2), this implies that on $\mathcal{B}$, we have $T_t(z)_{ab}\lesssim W^{-1}\eta_t^{-\frac{1}{2}}$. By the behavior of $S^{1/2}$ as an averaging operator, we have for all $a, b$ and $t\in [0, 1]$: $$(S^{1/2}T_t(z)S^{1/2})_{ab}\lesssim W^{-1}\eta_t^{-1/2}.$$ 
This means that by \hyperref[corollary-b5]{Corollary B.5} on the event $\mathcal{B}$, we get precisely: $$\frac{|G_t(z)_{ab}-m(z)\delta_{ab}|}{(S^{1/2}T_t(z)S^{1/2})_{ab}+S^{1/2}_{ab}+W^{-C}}\prec 1$$
Hence, $\mathbb{P}[\left\{\tau_{\sto, 2}\neq 1\}\cap \{\tau_{\sto, 1=1}\}\right]\rightarrow 0$ as $N\rightarrow \infty$, completing the theorem.

\subsection*{7.2 Theorem 2 - Quantum Diffusion}
\begin{boxtheorem}{Theorem 2: (Quantum Diffusion)}
For $|E|<2$ fixed, assume $\exists \nu>0$, s.t. $\eta \asymp W^2 N^{-2}$ and $W\geq W^{8/11+\nu}$. Then: $$\max_{x, y}|T_{xy}-\Theta_{xy}|\prec W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$$
\end{boxtheorem}
\textit{Proof} - By the previous theorem and the definition of stopping time, we know that $$\max_{a,b}|\E_1(z)_{ab}|\prec W^{-\frac{7}{4}} \eta_1^{-\frac{3}{2}}$$
But recall by our flow construction that $\eta_1=\eta$ and by extension, $T_1=T$ and $\Theta_1=\Theta$, meaning that: $$\max_{a,b}|\E_1(z)_{ab}| =\max_{x, y}|T_{xy}-\Theta_{xy}|\prec W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$$

@@

\noindent Now, let $\epsilon, D>0$ be arbitrary. Given that we work in the context of the delocalization, we have $W\gg N^{\frac{1}{2}+\beta} = N^{\frac{1+2\beta}{2}}\Rightarrow W^{-1/2}\ll N^{-\frac{1+2\beta}{4}}$ for any fixed $\beta>0$. Hence, let us choose $\delta \in \left(0, \frac{2\beta+1}{4\epsilon}\right)$ by the density of the rationals. By Def.1 of stochastic domination, from $(8.7)$, we have that: $$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)>N^{\delta\epsilon}(1+W^{-1/2}\Xi_N(s))\right)<N^{-D}$$
$$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)>N^{\delta\epsilon}+N^{\delta\epsilon} W^{-1/2}\Xi_N(s)\right)<N^{-D}$$
$$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)(1-N^{\delta\epsilon} W^{-1/2})>N^{\delta\epsilon}\right)<N^{-D}$$
$$\sup_{s\in S_N}\mathbb{P}\left(\Xi_N(s)>\frac{N^{\delta\epsilon}}{(1-N^{\delta\epsilon} W^{-1/2})}\right)<N^{-D}$$
By the delocalization scale and per our choice $\delta\epsilon <\frac{2\beta+1}{4\epsilon}$, we have that $$N^{\delta\epsilon}W^{-1/2} \ll N^{\delta\varepsilon} \cdot N^{-1/4-\beta/2} = N^{\delta\varepsilon-1/4-\beta/2}$$
$$1-N^{\delta\epsilon} W^{-1/2}\gg 1-N^{\delta\epsilon -\frac{1+2\beta}{4}}$$
$$\frac{N^{\delta\epsilon}}{1-N^{\delta\epsilon} W^{-1/2}}\ll \frac{N^{\delta\epsilon}}{1-N^{{\delta\epsilon}-\frac{1+2\beta}{4}}}$$

\noindent This means that $\exists N_0$, s.t. for all $N\geq N_0$, the latter event is a subset: $$ \left\{\Xi_N(s)>\frac{N^\epsilon}{1-N^{{\delta\epsilon}-\frac{1+2\beta}{4}}}\right\}\subset \left\{\Xi_N(s)>\frac{N^{\delta\epsilon}}{(1-N^{\delta\epsilon} W^{-1/2})}\right\}$$
$$\Rightarrow \mathbb{P} \left\{\Xi_N(s)>\frac{N^{\delta\epsilon}}{1-N^{\delta\epsilon-\frac{1+2\beta}{4}}}\right\}\leq\mathbb{P} \left\{\Xi_N(s)>\frac{N^{\delta\epsilon}}{(1-N^{\delta\epsilon} W^{-1/2})}\right\}\leq N^{-D},$$
by Boole's inequality. Now, this means that we, need to find $\gamma$ and $N_0'$, s.t. $\forall N\geq N_0'$:

$$N^\gamma\gg \frac{N^{\delta\epsilon}}{1-N^{\delta\epsilon-\frac{1+2\beta}{4}}}$$
